Partitioning mortality into growth-dependent and growth-independent hazards across 203 tropical tree species

### 2018-11-03 ms/manuscript.Rnw
proof edits

\author[a,1]{James S Camac}
\author[b]{Richard Condit}
\author[a,2]{Richard G FitzJohn}
\author[c]{Lachlan McCalman}
\author[c]{Daniel Steinberg}
\author[a]{Mark Westoby}
\author[d]{S Joseph Wright}
\author[e]{Daniel S Falster}

\affil[a]{Department of Biological Sciences, Macquarie University, Sydney, New South Wales, Australia}
\affil[b]{Field Museum of Natural History, Chicago, IL, USA and Morton Arboretum, Lisle, IL, USA}
\affil[c]{Data61, CSIRO, Australia}
\affil[d]{Smithsonian Tropical Research Institute, Apartado 0843–03092, Balboa, Panama}
\affil[e]{Evolution and Ecology Research Centre, School of Biological, Earth and Environmental Science, University of New South Wales, Sydney, New South Wales, Australia}

@@

% Please add here a significance statement to explain the relevance of your work
% 120 words
\significancestatement{We present a model that partitions rates of tropical tree mortality into growth-dependent and growth-independent hazards. This creates the opportunity to examine the relative contributions of within-species and across-species variation on tropical tree mortality rates, but also, how species traits affect each hazard. We parameterize this model using \textgreater 400,000 observed survival records collected over a 15-year period at Barro Colorado Island from more than 180,000 individuals across 203 species. We show that marginal carbon budgets are a major contributor to tree death on Barro Colorado Island. Moreover, we found that while species' light demand, maximum dbh and wood density affected tree mortality in different ways, they explained only a small fraction of the total variability observed among species.}

@@

\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: james.camac@gmail.com. Present address: Centre of Excellence for Biosecurity Risk Analysis, School of Biological Sciences, The University of Melbourne, Melbourne, Victoria, Australia

@@

% 250 words
\begin{abstract}
Tree death drives population dynamics, nutrient cycling, and evolution within plant communities. Mortality variation across species is thought to be influenced by different factors relative to variation within species. The unified model provided here separates mortality rates into growth-dependent and growth-independent hazards. This model creates the opportunity to simultaneously estimate the descriptors of variation both across and within species. Moreover, it provides the ability to examine how species traits affect growth-dependent and growth-independent hazards. We derive this unified mortality model  using cross-validated Bayesian methods coupled with mortality data collected over three census intervals for 203 tropical rainforest tree species at Barro Colorado Island (BCI) Panama. We found that growth independent mortality tended to be higher in species with lower wood density, higher light requirements and smaller maximum dbh. Mortality due to marginal carbon budget as measured by near-zero growth rate tended to be higher in species with lower wood density and higher light demand. The total mortality variation attributable to differences among species was large relative to variation explained by these traits, emphasizing that much remains to be understood. This additive hazards model strengthens our capacity to parse and understand individual-level mortality in highly diverse tropical forests, and hence to predict its consequences. 
\end{abstract}

@@

\begin{figure}[!hb]
\centering
\includegraphics{figures/fig1.pdf}
\caption{Outline of methodology.
A) We consider three alternative hazard functions: 1) a growth-independent baseline hazard (A, left), 2) a growth-dependent hazard, where growth rate is a proxy for carbon budget (A, middle); and 3) a function that combines growth-dependent with growth-independent hazards (A, right). The parameters of the models are biologically interpretable: $\gamma$ is the baseline or growth-independent hazard; $\alpha$ defines the low growth rate (i.e. very marginal carbon budget) effect on the overall hazard; $\beta$ captures the sensitivity or curvature of the response to declining growth rate. In the most complex model (Panel A; right), $\alpha$ and $\beta$ capture growth-dependent hazards, while $\gamma$ captures growth-independent hazards that kill a plant, irrespective of its growth rate/carbon budget (e.g. windfall, fire). For each model form, we consider two alternative measures of growth, $X$ (basal area and stem diameter growth), and we allow for species-level effects on the parameters $\alpha$, $\beta$ and $\gamma$.
B) Our data consist of repeat measures of stem diameter ($D$) and status ($S$, 0 = alive or 1 = dead) for individual trees at specific census dates ($t_1$, $t_2$, $t_3$).
C) Each model's skill in predicting observed outcomes ($S$) is quantified via the log-loss function (eq. \ref{eq:logloss}). Lower logloss means higher predictive accuracy.
D) The predictive skill of alternative models was evaluated via 10-fold cross validation. The entire dataset is split into 10 folds ($F_1$, ..., $F_{10}$). Alternative models were fit 10 times ($M_1$, ..., $M_{10}$), using different combinations of testing (1 fold; orange) and training (9 folds; green) data. Predictive capacity was assessed by averaging the log losses obtained from the 10 test data predictions.}
\label{fig1}
\end{figure}

To illustrate the proposed mortality model, we assess its ability to predict the survival status of \Sexpr{n_inds(BCI_training_full)} individual trees from \Sexpr{n_spp(BCI_training_full)} species, recorded over three 5-year census periods at Barro Colorado Island (BCI), Panama. Applying the additive model, we: i) estimate growth-dependent and growth-independent effects from long-term census data, ii) estimate species-level differences in these effects, and iii) assess how well species-level differences are predicted from three fundamental traits, previously shown to correlate with mortality: wood density, light demand, and maximum dbh \cite{king:2006va, Kraft:2010kq, Wright:2010fl, Ruger:2011cv}. These traits show only weak correlation with each other (r$^2$ \textless 0.13) and thus provide relatively independent measures of a species ecological strategy (SI Appendix, Fig. S7).

Using a likelihood-based penalty for incorrect prediction, we used 10-fold cross-validation to evaluate the skill of different models in their ability to predict outcomes in novel data (i.e. not used in model fitting; Fig. \ref{fig1}C-D; see Methods) \cite{Hooten:2015, Stone:1997tg}. In total \Sexpr{n_obs(BCI_training_full)} observations were used to fit these models. Evaluating models in this way is computationally expensive. Therefore rather than running all possible model formulations, we fitted models across six levels of increasing complexity, described below, and comparing their predictive skill. We then use the final model to assess the effects of species traits on each parameter and quantify their relative contributions on the overall model performance. Relative contributions of parameters were assessed using $r^2$ and Area Under the ROC Receiver Curve (AUROC). Whereby, $r^2$ estimates were derived from modelled predictions vs proportion of individual deaths observed between 1995 and 2010 for each species, and thus measured the amount of species-level variation explained by the model \cite{Visser_functional_2016}. By contrast, AUROC examined the model's ability to distinguish individuals that died from individuals that survived, irrespective of species.

@@

\begin{figure*}[!ht]
\includegraphics{figures/fig2.pdf}
\caption{Predictive skill of alternative hazard functions. Values are mean ($\pm$ 95\% confidence intervals) logarithmic loss, with lower values implying greater predictive skill and a logloss $\approx$ 0.693 indicating a model unable to distinguish an observed death from a survival.
A) Six sequential stages of model selection with each stage increasing in complexity. \textbf{Stage 1:} Baseline: constant ($\gamma$) vs Baseline with census effect ($\gamma\,\delta_t$); \textbf{Stage 2:} Best model of stage 1 compared to hazard functions that also account for individual growth rate (i.e. carbon budget). This includes two possible hazard functions: a growth-dependent only hazard ($(\alpha e^{-\beta X_{i}})\delta_t$) and a baseline + growth-dependent hazard with census effects ($(\alpha e^{-\beta X_{i}} + \gamma)\delta_t$); \textbf{Stage 3:} The effect of a species' maximum dbh ($\psi$), wood density ($\rho$) and light demand ($\upsilon$) on $\alpha$, $\beta$ and $\gamma$ of most predictive model from stage 2; \textbf{Stage 4:} 2-trait combinations; \textbf{Stage 5:} three-trait combination; \textbf{Stage 6:} Maximum predictive performance by accounting for all species differences via the inclusion of species random effects, $s$ on all three parameters. Symbols represent the three hazard functional form: baseline (circle), growth-dependent (triangle), and hazard containing both growth-independent and growth-dependent hazards (square). Colours represent models that don't account for individual growth (white); use individual basal area growth (grey); or use diameter increment growth (black).}
\label{fig2}
\end{figure*}

\subsection*{Stage 1: Mortality over time}
We first assess whether average mortality rates varied among the three censuses, in other words whether a census effect is justified in Eq \ref{eq:model}. Comparing a model with a constant mortality rate $\gamma$ to a model where the constant is scaled by a census effect $\gamma\delta_{t}$, and examining the three 5-year census intervals, average mortality rates declined over time. The highest proportion of death occurred from 1995\textendash 2000 (\Sexpr{prop_deaths(BCI_training_full, 1)}\%) followed by 2000\textendash 2005 (\Sexpr{prop_deaths(BCI_training_full, 2)}\%) and 2005\textendash 2010 (\Sexpr{prop_deaths(BCI_training_full, 3)}\%). Consequently, we observed a small increase in predictive skill when a census effect was included (leftmost two points under stage 1 in Fig. \ref{fig2}; see also Table.~\ref{table:1} for quantification of gain in predictive power under successive models).

\subsection*{Stage 2: Hazard functions}
Second, we assessed the predictive performance of three alternative hazard functions (Fig. \ref{fig1}A). The simplest form (Fig. \ref{fig1}A, left) assumes a constant hazard. The second form (Fig. \ref{fig1}A, middle) assumes the risk of dying declines towards an asymptote of zero as growth increases. The third "additive" model (Fig. \ref{fig1}BA right) is the sum of the previous two. In parallel, we investigated also which growth measure (increment of stem diameter or of stem area) predicts growth-dependent mortality better. Comparing the three hazard functions, the additive function significantly outperformed both the baseline (the best model in stage 1) and growth-dependent only hazard functions (right vs left of the four symbols under stage 2 in Fig. \ref{fig2}). Predictive skill was higher when using stem diameter growth compared to stem-area growth (Fig. \ref{fig2} stage 2, open vs closed symbols).

\subsection*{Stages 3-5: Species traits}
In stages 3-5, we take the additive model using dbh growth rate (the most predictive hazard function and growth rate measure from stage 2) and ask how addition of 1, 2 and 3 species-level traits improves model performance. Specifically, we introduce effects of the traits maximum dbh ($\psi$), light demand ($\upsilon$) and wood density ($\rho$) on the parameters $\alpha$, $\beta$ and $\gamma$ in Eq. \ref{eq:model}, at first one at a time (stage 3), then in 2-trait (stage 4) and 3-trait (stage 5) combinations. In general we found that including trait effects on these parameters decreased logarithmic loss, and thus, increased predictive accuracy (Fig.~\ref{fig2}). Allowing all three traits to have an effect on these parameters resulted in higher predictive capacity relative to models that only incorporated one or two of the traits.

@@

The findings of this study provide empirical support for dynamic vegetation models that predict tree mortality as a function of both growth-dependent and independent hazards \cite{Pacala:1996vm, Moorcroft:2001ws, McDowell:2011dr}. Regardless of growth measure (our proxy for carbon budget), incorporating both hazards significantly improved model predictive skill. This was because the growth-dependent hazard allows for deaths associated with low carbon budgets, and as a consequence, incorporates intra-specific variability attributed to carbon related stresses such as competition, parasites and herbivory. By contrast, the growth-independent hazard accounts for deaths caused by events that arise irrespective of an individual's carbon budget such as windthrow or lightning strike.

@@

\subsection*{Evaluating model skill}
Predictive skill was quantified by estimating the average log loss across 10-folds for held-out data, $\bar{\mathcal{L}}$ (Fig. \ref{fig1}). Logarithmic loss -- commonly known as log loss, $\mathcal{L}$, measures the skill of a model by penalizing incorrect predictions, based on how wrong the predicted probability is from the observed outcome, $S_i$ (Fig. \ref{fig1}C). Lower $\mathcal{L}$ implies greater skill. The average log loss across all individuals for the $k$th fold of held-out data, $\mathcal{L}_k$, is then
\begin{equation} \label{eq:logloss}
\mathcal{L}_k = - \frac1{N_k}\,\sum_{i=1}^{N_k}\left(S_i\,\log(p_{i,t_2\rightarrow t_3}) + ( 1 - S_i)\,\log(1 - p_{i,t_2\rightarrow t_3})\right),
\end{equation}
where $N_k$ refers to the number of observations in fold.
}

>>>>

\author[a,1]{James S Camac}
\author[b,c]{Richard Condit}
\author[a,2]{Richard G FitzJohn}
\author[d]{Lachlan McCalman}
\author[d]{Daniel Steinberg}
\author[a]{Mark Westoby}
\author[e]{S Joseph Wright}
\author[f]{Daniel S Falster}

\affil[a]{Department of Biological Sciences, Macquarie University, Sydney, NSW, 2109, Australia}
\affil[b]{Field Museum of Natural History, Chicago, IL, 60605, United States of America}
\affil[c]Morton Arboretum, Lisle, IL ,60532, United States of America}
\affil[d]{Data61, Commonwealth Scientific and Industrial Research Organisation (CSIRO), Sydney, NSW, 2015, Australia}
\affil[e]{Smithsonian Tropical Research Institute, Apartado 0843–03092, Balboa, Panama}
\affil[f]{Evolution and Ecology Research Centre, School of Biological, Earth and Environmental Sciences, University of New South Wales, Sydney, NSW, 2052, Australia}

@@

% Please add here a significance statement to explain the relevance of your work
% 120 words
\significancestatement{We present a model that partitions rates of tropical tree mortality into growth-dependent and growth-independent hazards. This creates the opportunity to examine the relative contributions of within-species and across-species variation on tropical tree mortality rates, but also, how species traits affect each hazard. We parameterize this model using \textgreater 400,000 observed survival records collected over a 15-year period at Barro Colorado Island from more than 180,000 individuals across 203 species. We show that marginal carbon budgets are a major contributor to tree death on Barro Colorado Island. Moreover, we found that while species' light demand, maximum diameter at breast height (dbh) and wood density affected tree mortality in different ways, they explained only a small fraction of the total variability observed among species.}

@@

\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: james.camac@gmail.com. Present address: Centre of Excellence for Biosecurity Risk Analysis, School of Biological Sciences, The University of Melbourne, Melbourne, VIC, 3010, Australia

@@

% 250 words
\begin{abstract}
Tree death drives population dynamics, nutrient cycling, and evolution within plant communities. Mortality variation across species is thought to be influenced by different factors relative to variation within species. The unified model provided here separates mortality rates into growth-dependent and growth-independent hazards. This model creates the opportunity to simultaneously estimate these hazards both across and within species. Moreover, it provides the ability to examine how species traits affect growth-dependent and growth-independent hazards. We derive this unified mortality model  using cross-validated Bayesian methods coupled with mortality data collected over three census intervals for 203 tropical rainforest tree species at Barro Colorado Island (BCI) Panama. We found that growth independent mortality tended to be higher in species with lower wood density, higher light requirements and smaller maximum diameter at breast height (dbh). Mortality due to marginal carbon budget as measured by near-zero growth rate tended to be higher in species with lower wood density and higher light demand. The total mortality variation attributable to differences among species was large relative to variation explained by these traits, emphasizing that much remains to be understood. This additive hazards model strengthens our capacity to parse and understand individual-level mortality in highly diverse tropical forests, and hence to predict its consequences. 
\end{abstract}

@@

\begin{figure}[!hb]
\centering
\includegraphics{figures/fig1.pdf}
\caption{Outline of methodology.
A) We consider three alternative hazard functions: 1) a growth-independent baseline hazard (A, left), 2) a growth-dependent hazard, where growth rate is a proxy for carbon budget (A, middle); and 3) a function that combines growth-dependent with growth-independent hazards (A, right). The parameters of the models are biologically interpretable: $\gamma$ is the baseline or growth-independent hazard; $\alpha$ defines the low growth rate (i.e. very marginal carbon budget) effect on the overall hazard; $\beta$ captures the sensitivity or curvature of the response to declining growth rate. In the most complex model (Panel A; right), $\alpha$ and $\beta$ capture growth-dependent hazards, while $\gamma$ captures growth-independent hazards that kill a plant, irrespective of its growth rate/carbon budget (e.g. windfall, fire). For each model form, we consider two alternative measures of growth, $X$ (basal area and stem diameter growth), and we allow for species-level effects on the parameters $\alpha$, $\beta$ and $\gamma$.
B) Our data consist of repeat measures of stem diameter ($D$) and status ($S$, 0 = alive or 1 = dead) for individual trees at specific census dates ($t_1$, $t_2$, $t_3$). Tree illustration courtesy of Tracey Saxby, Integration and Application Network, University of Maryland Center for Environmental Science (\url{ian.umces.edu/imagelibrary/}).
C) Each model's skill in predicting observed outcomes ($S$) is quantified via the log-loss function (eq. \ref{eq:logloss}). Lower logloss means higher predictive accuracy.
D) The predictive accuracy of alternative models was evaluated via 10-fold cross validation. The entire dataset is split into 10 folds ($F_1$, ..., $F_{10}$). Alternative models were fit 10 times ($M_1$, ..., $M_{10}$), using different combinations of testing (1 fold; orange) and training (9 folds; green) data. Predictive accuracy was assessed by averaging the log loss obtained from the 10 test data predictions.}
\label{fig1}
\end{figure}

To illustrate the proposed mortality model, we assess its ability to predict the survival status of \Sexpr{n_inds(BCI_training_full)} individual trees from \Sexpr{n_spp(BCI_training_full)} species, recorded over three 5-year census periods at Barro Colorado Island (BCI), Panama. Applying the additive model, we: i) estimate growth-dependent and growth-independent effects from long-term census data, ii) estimate species-level differences in these effects, and iii) assess how well species-level differences are predicted from three fundamental traits, previously shown to correlate with mortality: wood density, light demand, and maximum diameter at breast height (dbh) \cite{king:2006va, Kraft:2010kq, Wright:2010fl, Ruger:2011cv}. These traits show only weak correlation with each other (r$^2$ \textless 0.13) and thus provide relatively independent measures of a species ecological strategy (SI Appendix, Fig. S7).

Using a likelihood-based penalty for incorrect prediction, we used 10-fold cross-validation to evaluate the skill of different models in their ability to predict outcomes in novel data (i.e. not used in model fitting; Fig. \ref{fig1}C-D; see Methods) \cite{Hooten:2015, Stone:1997tg}. In total \Sexpr{n_obs(BCI_training_full)} observations were used to fit these models. Evaluating models in this way is computationally expensive. Therefore rather than running all possible model formulations, we fitted models across six levels of increasing complexity, described below, and comparing their predictive accuracy We then use the final model to assess the effects of species traits on each parameter and quantify their relative contributions on the overall model performance. Relative contributions of parameters were assessed using $r^2$ and Area Under the ROC Receiver Curve (AUROC). Whereby, $r^2$ estimates were derived from modelled predictions vs proportion of individual deaths observed between 1995 and 2010 for each species, and thus measured the amount of species-level variation explained by the model \cite{Visser_functional_2016}. By contrast, AUROC examined the model's ability to distinguish individuals that died from individuals that survived, irrespective of species.

@@

\begin{figure*}[!ht]
\includegraphics{figures/fig2.pdf}
\caption{Predictive accuracy of alternative hazard functions. Values are mean ($\pm$ 95\% confidence intervals) logarithmic loss, with lower values implying greater predictive accuracy and a logloss $\approx$ 0.693 indicating a model unable to distinguish an observed death from a survival.
A) Six sequential stages of model selection with each stage increasing in complexity. \textbf{Stage 1:} Baseline: constant ($\gamma$) vs Baseline with census effect ($\gamma\,\delta_t$); \textbf{Stage 2:} Best model of stage 1 compared to hazard functions that also account for individual growth rate (i.e. carbon budget). This includes two possible hazard functions: a growth-dependent only hazard ($(\alpha e^{-\beta X_{i}})\delta_t$) and a baseline + growth-dependent hazard with census effects ($(\alpha e^{-\beta X_{i}} + \gamma)\delta_t$); \textbf{Stage 3:} The effect of a species' maximum dbh ($\psi$), wood density ($\rho$) and light demand ($\upsilon$) on $\alpha$, $\beta$ and $\gamma$ of most predictive model from stage 2; \textbf{Stage 4:} 2-trait combinations; \textbf{Stage 5:} three-trait combination; \textbf{Stage 6:} Maximum predictive performance by accounting for all species differences via the inclusion of species random effects, $s$ on all three parameters. Symbols represent the three hazard functional form: baseline (circle), growth-dependent (triangle), and hazard containing both growth-independent and growth-dependent hazards (square). Colours represent models that don't account for individual growth (white); use individual basal area growth (grey); or use diameter increment growth (black).}
\label{fig2}
\end{figure*}

\subsection*{Stage 1: Mortality over time}
We first assess whether average mortality rates varied among the three censuses, in other words whether a census effect is justified in Eq \ref{eq:model}. Comparing a model with a constant mortality rate $\gamma$ to a model where the constant is scaled by a census effect $\gamma\delta_{t}$, and examining the three 5-year census intervals, average mortality rates declined over time. The highest proportion of death occurred from 1995\textendash 2000 (\Sexpr{prop_deaths(BCI_training_full, 1)}\%) followed by 2000\textendash 2005 (\Sexpr{prop_deaths(BCI_training_full, 2)}\%) and 2005\textendash 2010 (\Sexpr{prop_deaths(BCI_training_full, 3)}\%). Consequently, we observed a small increase in predictive accuracy when a census effect was included (leftmost two points under stage 1 in Fig. \ref{fig2}; see also Table.~\ref{table:1} for quantification of gain in predictive power under successive models).

\subsection*{Stage 2: Hazard functions}
Second, we assessed the predictive accuracy of three alternative hazard functions (Fig. \ref{fig1}A). The simplest form (Fig. \ref{fig1}A, left) assumes a constant hazard. The second form (Fig. \ref{fig1}A, middle) assumes the risk of dying declines towards an asymptote of zero as growth increases. The third "additive" model (Fig. \ref{fig1}BA right) is the sum of the previous two. In parallel, we investigated also which growth measure (increment of stem diameter or of stem area) predicts growth-dependent mortality better. Comparing the three hazard functions, the additive function significantly outperformed both the baseline (the best model in stage 1) and growth-dependent only hazard functions (right vs left of the four symbols under stage 2 in Fig. \ref{fig2}). Predictive accuracy was higher when using stem diameter growth compared to stem-area growth (Fig. \ref{fig2} stage 2, open vs closed symbols).

\subsection*{Stages 3-5: Species traits}
In stages 3-5, we take the additive model using dbh growth rate (the most predictive hazard function and growth rate measure from stage 2) and ask how addition of 1, 2 and 3 species-level traits improves model performance. Specifically, we introduce effects of the traits maximum dbh ($\psi$), light demand ($\upsilon$) and wood density ($\rho$) on the parameters $\alpha$, $\beta$ and $\gamma$ in Eq. \ref{eq:model}, at first one at a time (stage 3), then in 2-trait (stage 4) and 3-trait (stage 5) combinations. In general we found that including trait effects on these parameters decreased logarithmic loss, and thus, increased predictive accuracy (Fig.~\ref{fig2}). Allowing all three traits to have an effect on these parameters resulted in higher predictive accuracy relative to models that only incorporated one or two of the traits.

@@

The findings of this study provide empirical support for dynamic vegetation models that predict tree mortality as a function of both growth-dependent and independent hazards \cite{Pacala:1996vm, Moorcroft:2001ws, McDowell:2011dr}. Regardless of growth measure (our proxy for carbon budget), incorporating both hazards significantly improved model predictive accuracy. This was because the growth-dependent hazard allows for deaths associated with low carbon budgets, and as a consequence, incorporates intra-specific variability attributed to carbon related stresses such as competition, parasites and herbivory. By contrast, the growth-independent hazard accounts for deaths caused by events that arise irrespective of an individual's carbon budget such as windthrow or lightning strike.

@@

\subsection*{Evaluating model predictive accuracy}
Predictive skill was quantified by estimating the average log loss across 10-folds for held-out data, $\bar{\mathcal{L}}$ (Fig. \ref{fig1}). Logarithmic loss -- commonly known as log loss, $\mathcal{L}$, measures the skill of a model by penalizing incorrect predictions, based on how wrong the predicted probability is from the observed outcome, $S_i$ (Fig. \ref{fig1}C). Lower $\mathcal{L}$ implies greater skill. The average log loss across all individuals for the $k$th fold of held-out data, $\mathcal{L}_k$, is then
\begin{equation} \label{eq:logloss}
\mathcal{L}_k = - \frac1{N_k}\,\sum_{i=1}^{N_k}\left(S_i\,\log(p_{i,t_2\rightarrow t_3}) + ( 1 - S_i)\,\log(1 - p_{i,t_2\rightarrow t_3})\right),
\end{equation}
where $N_k$ refers to the number of observations in fold.
}
